Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling

### 2021-04-07 manuscript/manuscript.Rmd
Fix formatting and bugs

---
title: Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - categorical data
  - neural networks
  - embeddings
  - attention
  - severity Modeling
  - National Flood Insurance Program 
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
    \newcommand{\norm}[1]{\| #1 \|}
---

@@

Categorical data are modelled by actuaries in many different contexts, from pricing of insurance products to reserving for the liabilities generated by these products. Sometimes, these data are modelled in an explicit manner, for example, when building models that apply across multiple categories, a form of dummy coding is usually used. For example, when building models to price the frequency of motor insurance claims, claim experience relating to different types of motor vehicle will often be modelled by including a (single) factor within models that modifies the relative frequency predicted for each type of vehicle. In other cases, Modeling is performed for each category separately, thus the categorical data is used within the Modeling in an implicit manner, for example, common practice is to estimate reserves for different lines of business separately, meaning to say, with no parameters being shared across each of the reserving models.

@@

This approach to Modeling categorical data has several advantages over more traditional treatments of categorical data. Using categorical embeddings instead of traditional techniques has been shown to increase predictive accuracy of models, for example, see @Richman2018 in the context of pricing. Models incorporating categorical embeddings can be pre-calibrated to traditional actuarial models, increasing the speed with which these models can be calibrated and leading to models with better explainability [@wuthrich2019yes]. Finally, the similarity between the vectors learned for different categories can be inspected, sometimes leading to insights into the workings of models, see, for example, @Kuo2019.

On the other hand, several open questions about the use of embeddings within actuarial work remain, which we aim to address in this study. First, hyperparameter settings for embeddings, such as the dimensions of the embedding layer and the use of regularization techniques such as dropout or normalization, that achieve optimal predictive performance has not yet been studied in detail in the actuarial literature. In this work, we aim to study how embeddings using different settings perform in the context of a large-scale predictive Modeling problem, and give guidance on the process that can be followed to determine this in other problems. Although neural network have been shown to achieve excellent predictive accuracy on actuarial tasks, many actuaries still prefer to use GLM models for pricing tasks, thus, the issue of whether transferring embeddings to GLM models can achieve better performance is considered in this paper. <!-- Traditional actuarial techniques such as credibility theory have been used to work with some types of categorical data, but no study has been --> <!-- performed whether this can be applied within embeddings; here, we investigate whether performance is enhanced by applying credibility theory on embeddings 
relating to categorical variables with many labels. Whereas embedding layers are usually considered in the context of categorical data,the 
option exists to quantize numerical data and model it using embeddings, however, the results of doing so have not been investigated. --> Finally, in the past several years, a new type of neural network architecture based on attention [@Vaswani2017] has been successfully used on embeddings in the field of natural language processing and we incorporate attention based models into our predictive Modeling example to demonstrate the use of these models for actuarial purposes.

@@

The rest of this manuscript is organized as follows. Section \ref{lit_review} reviews recent applications of embeddings in the actuarial literature. Section \ref{defns} provides the notation used in the paper and defines GLMs, neural networks and related Modeling concepts, including embeddings and attention. In Section \ref{Modeling}, we provide initial models for the NFIP dataset and consider the influence of hyperparameter choices on the results of the neural network model. That section also considers how successfully the embeddings used in the neural network model can be transferred to a GLM model of the same data and how learned embeddings can be interpreted. We focus on attention based models in Section \ref{attention_models}. Finally, Section \ref{Conclusions} provides a discussion of the results of this paper and considers avenues for future research.

@@

A different approach to Modeling categorical data is credibility theory (see @Buhlmann2006 for an overview), which, in the context of rating, can be applied to derive premiums that reflect the experience of a particular policyholder, by estimating premiums as a weighted average between the premium produced using the collective experience (i.e. of all policyholders) and the premium produced using the experience of the particular policyholder. The weight used in this average is called a credibility factor and is calculated with reference to te variability of the policyholder experience relative to the variability of the group experience. In this context, the implicit categorical variable is the policyholder under consideration.

Generalized Linear Mixed Models (GLMMs) are an extension of GLMs that are designed for Modeling categorical data using a principle very similar to that of credibility theory [@Klinker2010]. Instead of calibrating indicator variables for each level of the category, GLMMs estimate effects for each of these levels as a combination of the overall group mean and the experience in each level of the category.

Embedding layers represent a different approach to the problem of Modeling categorical data that was recently introduced in an actuarial context. Note that in the next section, we reflect on similarities between the conventional approaches discussed above and embedding layers. @Richman2018 reviewed the concept of embedding layers and connected the sharing of information across categories to the familiar concept of credibility theory. In that work, two applications of embedding layers were demonstrated. The first of these was in a Property and Casualty (P&C) pricing context, it was shown that the out-of-sample accuracy of a neural network trained to predict claims frequencies on motor third party liability was enhanced by Modeling the categorical variables within this dataset using embedding layers. Second, a neural network with embedding layers was used to model all of the mortality rates in the Human Mortality Database, where the differences in population mortality across countries and the differences in mortality at different ages were modelled with embedding layers, again producing more accurate out of sample performance than the other models tested.

@@

In this study, we are concerned with regression Modeling, which is the task of predicting an unknown outcome $y$ on the basis of information about that outcome contained in predictor variables, or features, stored in a matrix $X$. For simplicity, we only consider the case of univariate outcomes, i.e., $y \in \mathbb{R}^1$. The outcomes and the rows of the predictor variable matrix are indexed by $i \in \{1 \dots I\}$, where $i$ represents a particular observation of $(y_i, \mathbf{x}_i)$, where bold indicates that we are now dealing with a vector. The columns of the predictor variables are indexed by $j \in \{1 \dots J\}$, where $j$ represents a particular predictor variable, of which $J$ have been observed, thus, we use the notation $X_j$ to represent the $j$th predictor variable and $X \in \mathbb{R}^J$. Formally, we look to build regression models that map from the predictor variables $\mathbf{x}_i$ to the outcome $y$ using a function $f$ of the form:

@@

```{=tex}
\begin{equation} 
\label{NN}
\begin{split}
z^1 &= \sigma(a_1.X+ b_1) \\
z^2 &= \sigma(a_2.z^1+ b_2) \\
\vdots \\
z^K &= \sigma(a_{K}.z^{K-1}+ b_{K}) \\
\hat{y} &= \sigma(a_{K+1}.z^{K}+ b_{K+1}),
\end{split}
\end{equation}
```
where the regression parameters (weights) for each layer $k \in [1;K+1]$ are represented by the matrices $a_k$ and the intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the hyperbolic tangent (tanh) function or the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') as follows. First a loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ and the predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation). Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

@@

```{=tex}
\begin{equation} 
\label{embed_map}
  z_\mathcal{P}^j : \mathcal{P}^j \to \mathbb{R}^{q_\mathcal{P}^j }, \quad \quad p^j \mapsto  z_\mathcal{P}^j(p),
\end{equation}
```
where $q_\mathcal{P}^j$ is the dimension of the embedding for the $j$th categorical variable and $z_\mathcal{P}^j(.)$ is an implicit function that maps from the particular element of the labels $p$ to the embedding space. Equation \ref{embed_map} states that an embedding maps a level of a categorical variable to a numerical vector. This function is left implicit, meaning to say, we allow the embeddings to be derived during the process of fitting the model and do not attempt to specify exactly how the embeddings can be derived from the input data. In Table \@ref(tab:embed) we show an example of two dimensional embeddings for the state variables, where these have been generated randomly.

@@

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular Modeling tasks. Attention mechanisms allow deep neural nets to weight the covariates entering the model in a flexible manner; these types of applications in natural language processing are demonstrated in, for example, @Vaswani2017a. In an actuarial context, attention mechanisms can be understood as giving greater weight to covariates that are important for the Modeling task. In this section we provide a theoretical introduction to attention mechanisms and then discuss applications of attention to modeling tabular data. In the following sections of this paper, we then consider whether adding attention to an model that uses embeddings can provide better predictive accuracy. The following is intended as a high level review of some of the relevant theory behind attention based approaches; for more technical details the sources quoted in this section can be consulted.

@@

```{=latex}
\begin{equation}
X^*: \mathbb{R}^{(q \times d) \times (q \times d) \times (q \times d)} \to \mathbb{R}^{(q \times d)}, ~~
(Q, K, X') ~\mapsto ~ X^* = Attn(Q, K, X'),
\label{eq:attention}
\end{equation}
```
where $Q \in \mathbb{R}^{q \times d}$ is a matrix of query vectors, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors, $X' \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $X^* \in \mathbb{R}^{q \times d}$, the output of the attention mechanism, is a new matrix of values that have been reweighted according to their importance for the Modeling problem. As noted above, several different options for the mapping in Equation \ref{eq:attention} exist in the literature; here we describe the scaled dot-product self-attention mechanism used in @Vaswani2017a. In more detail, consider that $X'$ is a matrix of covariates relating to a regression problem, which, in the case of tabular data, will usually be composed of embeddings of the same dimension for several different categorical or numerical variables. In the case discussed here we have $q$ covariates mapped to embeddings of dimension $d$. For clarity, note that in Equation \ref{eq:embed_to_nn} we have used the symbol $X'$ to denote the vector of embeddings and numerical inputs which is passed into the neural network. We slightly overload this notation and, in what follows, we now use the symbol $X'$ to refer to the matrix of embeddings of dimension $q \times d$.

@@

An intuitive explanation of the attention formula is as follows: we start by considering the importance of all of the covariates that we have for our Modeling problem in the context of each individual covariate. This is done by assigning a score between zero and one to each covariate in the Modeling problem, and we repeat this process for each of the covariates. Then, we replace each of the original covariates with a new covariate that has been formed by taking a weighted average of the covariates and the scores.

@@

### Attention Modeling for Tabular Data

Several different approaches to applying attention within the Modeling of tabular data have been proposed in the literature. Perhaps the simplest way of doing this is to insert an attention layer between the embeddings and the rest of the neural network. Mathematically, we can modify Equation \ref{eq:embed_to_nn} as follows:

@@

Finally, a different way of incorporating attention into the modeling of tabular data has been proposed by @Ark2019, who refer to their model as TabNet. In summary, the main idea of the TabNet model is to try emulate the excellent performance of decision trees on tabular data Modeling using deep neural networks. This is done by structuring the neural networks to emulate a key aspect of decision tree Modeling, which is the efficient selection of relevant covariates for the model task. Within TabNet, an attention based model is used to select features by estimating attention scores for each covariate within $X$. The covariates are multiplied by the attention scores to down-weight less relevant features for the prediction problem. Other less important details of the TabNet model are given in @Ark2019.

# Predictive Modeling with Embeddings {#Modeling}

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is made available by the OpenFEMA initiative of the Federal Emergency Management Agency (FEMA)[^1]. Two datasets are made available by OpenFEMA: A policies dataset with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to data from 2000 to August 2019 and further subsample 100K data points to make experiments feasible on a CPU. The dataset includes a rich variety of variables, from occupancy type to flood zone. For our models, we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, which we list in Table \@ref(tab:variables)

@@

We note that the focus of our contribution is on applications of categorical embeddings in predictive modeling, rather than on fine-tuning models in order to beat or establish a benchmark. To this end, we do not perform systematic hyperparameter tuning in our experiments, and the performance metrics are provided to show that reasonable models utilizing categorical embeddings. However, as we describe the modeling procedure and results, we provide commentary on recommended practices in practical applications, especially with respect to the neural network models.

@@

```{=tex}
\begin{equation}
1 + |J^{\text{num}}| + \left(\sum_{j \in J^{\text{cat}}} n_{\mathcal{P}}^j\right) - |J^{\text{cat}}|:
(\#eq:glm1)
\end{equation}
```
one for the intercept, one for each numeric variable, and the numbers of levels minus one (due to dummy coding) for each categorical variable.

In Table \@ref(tab:relativities), we exhibit an exerpt of the relativities, or exponentiated fitted coefficients, of a couple categorical variables. Recall that the choice of the log link gives a multiplicative structure, so the interpretation of 1.34 for `occupancy_type = "Non-residential building"` is that, all else equal, the expected loss for a policy for a non-residential building is 1.34 times that of a policy for a single family home, which is the base, or reference, level for the factor.

@@

With a single hidden layer, this architecture is relatively simple; without the hidden layer and corresponding ReLU activation, we would actually recover a logistic regression structure. The choice of one for the embedding dimension is also due to simplicity: it is easily interpretable as the representation of each factor level as a point on the real number line; also, the trained embeddings are easily incorporated into a GLM which we will see in Model 3. In practice, the embedding dimension is a hyperparameter one can tune, for example via cross validation, and can vary for each variable. However, we also note that the choice of one for the embedding output dimension is not an unreasonable one, especially when the cardinalities of the categorical variables are not large (i.e., when they are in the 10s).

@@

```{=tex}
\begin{equation}
1 + |J^{\text{num}}| + |J^{\text{cat}}|.
(\#eq:glm2)
\end{equation}```
Hence, the model can be written as

```{=tex}
\begin{equation}
\mu_i = \beta_0 + \sum_{j \in J^{\text{num}}}\beta_{j} x_{i,j} + \sum_{j \in J^{\text{cat}}} \beta_{j} z_{i,j},
\end{equation}
```
where $z_{i,j}$ represents the embedding value for $i$th instance of the categorical column $X_j$.

@@

```{=tex}
\begin{equation}
p_{ij} = \frac{p_{j|i} + p_{i|j}}{2n}.
\end{equation}
```
On the other hand, in the target space, t-SNE assumes the neighbor selection is driven by a Student t-distribution with one degree of freedom, i.e., a Cauchy distribution. The similarity metric for a pair of points in $\mathcal{Y}$ is then defined as the joint probability

```{=tex}
\begin{equation}
q_{ij} = \frac{(1 + \norm{y_i - y_j}^2)^{-1}}{\sum_{k \neq l} (1 + \norm{y_k - y_l}^2)^{-1}}.
\end{equation}
```
With the similarity metrics defined for each space, we can then state the objective for "closeness" for the sets of points, which is the Kullback--Leibler divergence from Q to P:

```{=tex}
\begin{equation}
D_{KL}(P || Q) = \sum_{i \neq j} p_{ij} \log \frac{p_{ij}}{q_{ij}}. (\#eq:tsne-obj)
\end{equation}
```
We now circle back to Equation \@ref(eq:pji) and discuss $\sigma_i$. The variance of the Gaussian is allowed to vary depending on where the point is in the space; specifically, it should reflect the density or sparsness of surrounding points. To train this parameter, t-SNE introduces a hyperparameter, *perplexity*, which is a scalar value set by the user, and is defined as

```{=tex}
\begin{equation}
Perp(P_i) = 2^{H(P_i)},
\end{equation}```
where $H(P_i)$ denotes the entropy of $P_i$,

```{=tex}
\begin{equation}
H(P_i) = -\sum_j p_{j|i}\log_2 p_{j|i},
\end{equation}
```
and $P_i$ is the Gaussian parameterized by $x_i$ and $\sigma_i$. During training, t-SNE looks for $P_i$ that have perpelxity corresponding to the user input. Finally, the objective as stated in Equation \@ref(eq:tsne-obj) is differentiable, and is trained via gradient descent.

While t-SNE is a powerful visualization technique, there are a few points the modeler should be mindful of in practice. We refer the reader to @wattenberg2016how for a more thorough discussion, but stress here that the resulting plot can be very sensitive to the perplexity hyperparameter. In addition, for our use case, which is to explore the learned embeddings, we have fewer data points than most other applications in the literature. Accordingly, our reaosnable range of hyperparameter values need to be adjusted down. In Figures \@ref(fig:flood-zone-tsne), we display the t-SNE of the learned embeddings for `flood_zone` in Model 3 with various perplexity settings: 2, 3, 5, and 10; the learning rate and the number of steps are fixed at 100 and 10,000, respectively, for all plots. In the first three plots, we can see the semblance of clusters forming. In particular, flood zone D appears in its own cluster, which implies it is far away from other levels in the embedding space. When the perplexity is set to 10 (and beyond), however, we start to lose the geometric properties.

@@

Compared to the models presented in the previous section, the models presented here are closer to state of the art models that might be used for tabular data Modeling problems. To the best of our knowledge, this is the first analysis of these architectures within the actuarial literature.

@@

The final model we consider in this subsection is the TabNet architecture. The structure of the model is quite similar to Model 6, except that a more specialized form of self-attention is used to select the most important covariates for the Modeling problem.

>>>>

---
title: Categorical Predictor Embeddings and Attention Based Neural Networks in Actuarial Modeling
authors:
  - name: Kevin Kuo 
    affiliation: RStudio and Kasa AI
    email: kevin@kasa.ai
  - name: Ronald Richman
    affiliation: QED Actuaries and Consultants
    email: ronaldrichman@gmail.com
abstract: |
  Enter the text of your abstract here.
keywords:
  - categorical data
  - neural networks
  - embeddings
  - attention
  - severity modeling
  - National Flood Insurance Program 
bibliography: references.bib
output:
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
header-includes: |
    \usepackage{amsmath}
    \usepackage{mathtools}
    \usepackage{bm}
    \usepackage{bbm}
    \usepackage{float}
    \DeclarePairedDelimiter\ceil{\lceil}{\rceil}
    \newcommand{\norm}[1]{\| #1 \|}
---

@@

Categorical data are modelled by actuaries in many different contexts, from pricing of insurance products to reserving for the liabilities generated by these products. Sometimes, these data are modelled in an explicit manner, for example, when building models that apply across multiple categories, a form of dummy coding is usually used. For example, when building models to price the frequency of motor insurance claims, claim experience relating to different types of motor vehicle will often be modelled by including a (single) factor within models that modifies the relative frequency predicted for each type of vehicle. In other cases, modeling is performed for each category separately, thus the categorical data is used within the modeling in an implicit manner, for example, common practice is to estimate reserves for different lines of business separately, meaning to say, with no parameters being shared across each of the reserving models.

@@

This approach to modeling categorical data has several advantages over more traditional treatments of categorical data. Using categorical embeddings instead of traditional techniques has been shown to increase predictive accuracy of models, for example, see @Richman2018 in the context of pricing. Models incorporating categorical embeddings can be pre-calibrated to traditional actuarial models, increasing the speed with which these models can be calibrated and leading to models with better explainability [@wuthrich2019yes]. Finally, the similarity between the vectors learned for different categories can be inspected, sometimes leading to insights into the workings of models, see, for example, @Kuo2019.

On the other hand, several open questions about the use of embeddings within actuarial work remain, which we aim to address in this study. First, hyperparameter settings for embeddings, such as the dimensions of the embedding layer and the use of regularization techniques such as dropout or normalization, that achieve optimal predictive performance has not yet been studied in detail in the actuarial literature. In this work, we aim to study how embeddings using different settings perform in the context of a large-scale predictive modeling problem, and give guidance on the process that can be followed to determine this in other problems. Although neural network have been shown to achieve excellent predictive accuracy on actuarial tasks, many actuaries still prefer to use GLM models for pricing tasks, thus, the issue of whether transferring embeddings to GLM models can achieve better performance is considered in this paper. <!-- Traditional actuarial techniques such as credibility theory have been used to work with some types of categorical data, but no study has been --> <!-- performed whether this can be applied within embeddings; here, we investigate whether performance is enhanced by applying credibility theory on embeddings 
relating to categorical variables with many labels. Whereas embedding layers are usually considered in the context of categorical data,the 
option exists to quantize numerical data and model it using embeddings, however, the results of doing so have not been investigated. --> Finally, in the past several years, a new type of neural network architecture based on attention [@Vaswani2017] has been successfully used on embeddings in the field of natural language processing and we incorporate attention based models into our predictive modeling example to demonstrate the use of these models for actuarial purposes.

@@

The rest of this manuscript is organized as follows. Section \ref{lit_review} reviews recent applications of embeddings in the actuarial literature. Section \ref{defns} provides the notation used in the paper and defines GLMs, neural networks and related modeling concepts, including embeddings and attention. In Section \ref{modeling}, we provide initial models for the NFIP dataset and consider the influence of hyperparameter choices on the results of the neural network model. That section also considers how successfully the embeddings used in the neural network model can be transferred to a GLM model of the same data and how learned embeddings can be interpreted. We focus on attention based models in Section \ref{attention_models}. Finally, Section \ref{Conclusions} provides a discussion of the results of this paper and considers avenues for future research.

@@

A different approach to modeling categorical data is credibility theory (see @Buhlmann2006 for an overview), which, in the context of rating, can be applied to derive premiums that reflect the experience of a particular policyholder, by estimating premiums as a weighted average between the premium produced using the collective experience (i.e. of all policyholders) and the premium produced using the experience of the particular policyholder. The weight used in this average is called a credibility factor and is calculated with reference to te variability of the policyholder experience relative to the variability of the group experience. In this context, the implicit categorical variable is the policyholder under consideration.

Generalized Linear Mixed Models (GLMMs) are an extension of GLMs that are designed for modeling categorical data using a principle very similar to that of credibility theory [@Klinker2010]. Instead of calibrating indicator variables for each level of the category, GLMMs estimate effects for each of these levels as a combination of the overall group mean and the experience in each level of the category.

Embedding layers represent a different approach to the problem of modeling categorical data that was recently introduced in an actuarial context. Note that in the next section, we reflect on similarities between the conventional approaches discussed above and embedding layers. @Richman2018 reviewed the concept of embedding layers and connected the sharing of information across categories to the familiar concept of credibility theory. In that work, two applications of embedding layers were demonstrated. The first of these was in a Property and Casualty (P&C) pricing context, it was shown that the out-of-sample accuracy of a neural network trained to predict claims frequencies on motor third party liability was enhanced by modeling the categorical variables within this dataset using embedding layers. Second, a neural network with embedding layers was used to model all of the mortality rates in the Human Mortality Database, where the differences in population mortality across countries and the differences in mortality at different ages were modelled with embedding layers, again producing more accurate out of sample performance than the other models tested.

@@

In this study, we are concerned with regression modeling, which is the task of predicting an unknown outcome $y$ on the basis of information about that outcome contained in predictor variables, or features, stored in a matrix $X$. For simplicity, we only consider the case of univariate outcomes, i.e., $y \in \mathbb{R}^1$. The outcomes and the rows of the predictor variable matrix are indexed by $i \in \{1 \dots I\}$, where $i$ represents a particular observation of $(y_i, \mathbf{x}_i)$, where bold indicates that we are now dealing with a vector. The columns of the predictor variables are indexed by $j \in \{1 \dots J\}$, where $j$ represents a particular predictor variable, of which $J$ have been observed, thus, we use the notation $X_j$ to represent the $j$th predictor variable and $X \in \mathbb{R}^J$. Formally, we look to build regression models that map from the predictor variables $\mathbf{x}_i$ to the outcome $y$ using a function $f$ of the form:

@@

\begin{equation} 
\label{NN}
\begin{split}
z^1 &= \sigma(a_1.X+ b_1) \\
z^2 &= \sigma(a_2.z^1+ b_2) \\
\vdots \\
z^K &= \sigma(a_{K}.z^{K-1}+ b_{K}) \\
\hat{y} &= \sigma(a_{K+1}.z^{K}+ b_{K+1}),
\end{split}
\end{equation}

where the regression parameters (weights) for each layer $k \in [1;K+1]$ are represented by the matrices $a_k$ and the intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the hyperbolic tangent (tanh) function or the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') as follows. First a loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ and the predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation). Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

@@

\begin{equation} 
\label{embed_map}
  z_\mathcal{P}^j : \mathcal{P}^j \to \mathbb{R}^{q_\mathcal{P}^j }, \quad \quad p^j \mapsto  z_\mathcal{P}^j(p),
\end{equation}

where $q_\mathcal{P}^j$ is the dimension of the embedding for the $j$th categorical variable and $z_\mathcal{P}^j(.)$ is an implicit function that maps from the particular element of the labels $p$ to the embedding space. Equation \ref{embed_map} states that an embedding maps a level of a categorical variable to a numerical vector. This function is left implicit, meaning to say, we allow the embeddings to be derived during the process of fitting the model and do not attempt to specify exactly how the embeddings can be derived from the input data. In Table \@ref(tab:embed) we show an example of two dimensional embeddings for the state variables, where these have been generated randomly.

@@

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular modeling tasks. Attention mechanisms allow deep neural nets to weight the covariates entering the model in a flexible manner; these types of applications in natural language processing are demonstrated in, for example, @Vaswani2017a. In an actuarial context, attention mechanisms can be understood as giving greater weight to covariates that are important for the modeling task. In this section we provide a theoretical introduction to attention mechanisms and then discuss applications of attention to modeling tabular data. In the following sections of this paper, we then consider whether adding attention to an model that uses embeddings can provide better predictive accuracy. The following is intended as a high level review of some of the relevant theory behind attention based approaches; for more technical details the sources quoted in this section can be consulted.

@@

```{=latex}
\begin{equation}
X^*: \mathbb{R}^{(q \times d) \times (q \times d) \times (q \times d)} \to \mathbb{R}^{(q \times d)}, ~~
(Q, K, X') ~\mapsto ~ X^* = Attn(Q, K, X'),
\label{eq:attention}
\end{equation}
```
where $Q \in \mathbb{R}^{q \times d}$ is a matrix of query vectors, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors, $X' \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $X^* \in \mathbb{R}^{q \times d}$, the output of the attention mechanism, is a new matrix of values that have been reweighted according to their importance for the modeling problem. As noted above, several different options for the mapping in Equation \ref{eq:attention} exist in the literature; here we describe the scaled dot-product self-attention mechanism used in @Vaswani2017a. In more detail, consider that $X'$ is a matrix of covariates relating to a regression problem, which, in the case of tabular data, will usually be composed of embeddings of the same dimension for several different categorical or numerical variables. In the case discussed here we have $q$ covariates mapped to embeddings of dimension $d$. For clarity, note that in Equation \ref{eq:embed_to_nn} we have used the symbol $X'$ to denote the vector of embeddings and numerical inputs which is passed into the neural network. We slightly overload this notation and, in what follows, we now use the symbol $X'$ to refer to the matrix of embeddings of dimension $q \times d$.

@@

An intuitive explanation of the attention formula is as follows: we start by considering the importance of all of the covariates that we have for our modeling problem in the context of each individual covariate. This is done by assigning a score between zero and one to each covariate in the modeling problem, and we repeat this process for each of the covariates. Then, we replace each of the original covariates with a new covariate that has been formed by taking a weighted average of the covariates and the scores.

@@

### Attention modeling for Tabular Data

Several different approaches to applying attention within the modeling of tabular data have been proposed in the literature. Perhaps the simplest way of doing this is to insert an attention layer between the embeddings and the rest of the neural network. Mathematically, we can modify Equation \ref{eq:embed_to_nn} as follows:

@@

Finally, a different way of incorporating attention into the modeling of tabular data has been proposed by @Ark2019, who refer to their model as TabNet. In summary, the main idea of the TabNet model is to try emulate the excellent performance of decision trees on tabular data modeling using deep neural networks. This is done by structuring the neural networks to emulate a key aspect of decision tree modeling, which is the efficient selection of relevant covariates for the model task. Within TabNet, an attention based model is used to select features by estimating attention scores for each covariate within $X$. The covariates are multiplied by the attention scores to down-weight less relevant features for the prediction problem. Other less important details of the TabNet model are given in @Ark2019.

# Predictive Modeling with Embeddings {#modeling}

@@

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is made available by the OpenFEMA initiative of the Federal Emergency Management Agency (FEMA)[^1]. Two datasets are made available by OpenFEMA: A policies dataset with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the complete dataset contains over two million transactions, for the purposes of our experiments, we limit ourselves to data from 2000 to 2019 and further subsample 100K data points to make experiments feasible on a CPU. The dataset includes a rich variety of variables, from occupancy type to flood zone. For our models, we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, which we list in Table \@ref(tab:variables)

@@

We note that the focus of our contribution is on applications of categorical embeddings in predictive modeling, rather than on fine-tuning models in order to beat or establish a benchmark. To this end, we do not perform systematic hyperparameter tuning in our experiments, and the performance metrics are provided to show that reasonable models utilizing categorical embeddings can be developed (note in the next section, the models achieve significantly better performance). However, as we describe the modeling procedure and results, we provide commentary on recommended practices in practical applications, especially with respect to the neural network models.

@@

\begin{equation}
1 + |J^{\text{num}}| + \left(\sum_{j \in J^{\text{cat}}} n_{\mathcal{P}}^j\right) - |J^{\text{cat}}|:
(\#eq:glm1)
\end{equation}

one for the intercept, one for each numeric variable, and the numbers of levels minus one (due to dummy coding) for each categorical variable.

In Table \@ref(tab:relativities), we exhibit an exerpt of the relativities, or exponentiated fitted coefficients, of a couple of the categorical variables. Recall that the choice of the log link gives a multiplicative structure, so the interpretation of 1.34 for `occupancy_type = "Non-residential building"` is that, all else equal, the expected loss for a policy for a non-residential building is 1.34 times that of a policy for a single family home, which is the base, or reference, level for the factor.

@@

With a single hidden layer, this architecture is relatively simple; without the hidden layer and corresponding ReLU activation, we would actually recover a type of logistic regression structure. The choice of one for the embedding dimension is also due to simplicity: it is easily interpretable as the representation of each factor level as a point on the real number line; also, the trained embeddings are easily incorporated into a GLM which we will see in Model 3. In practice, the embedding dimension is a hyperparameter one can tune, for example via cross validation, and can vary for each variable. However, we also note that the choice of one for the embedding output dimension is not an unreasonable one, especially when the cardinalities of the categorical variables are not large (i.e., when they are in the 10s).

@@

\begin{equation}
1 + |J^{\text{num}}| + |J^{\text{cat}}|.
(\#eq:glm2)
\end{equation}
Hence, the model can be written as

\begin{equation}
\mu_i = \beta_0 + \sum_{j \in J^{\text{num}}}\beta_{j} x_{i,j} + \sum_{j \in J^{\text{cat}}} \beta_{j} z_{i,j},
\end{equation}

where $z_{i,j}$ represents the embedding value for $i$th instance of the categorical column $X_j$.

@@

\begin{equation}
p_{ij} = \frac{p_{j|i} + p_{i|j}}{2n}.
\end{equation}

On the other hand, in the target space, t-SNE assumes the neighbor selection is driven by a Student t-distribution with one degree of freedom, i.e., a Cauchy distribution. The similarity metric for a pair of points in $\mathcal{Y}$ is then defined as the joint probability

\begin{equation}
q_{ij} = \frac{(1 + \norm{y_i - y_j}^2)^{-1}}{\sum_{k \neq l} (1 + \norm{y_k - y_l}^2)^{-1}}.
\end{equation}

With the similarity metrics defined for each space, we can then state the objective for "closeness" for the sets of points, which is the Kullback--Leibler divergence from Q to P:

\begin{equation}
D_{KL}(P || Q) = \sum_{i \neq j} p_{ij} \log \frac{p_{ij}}{q_{ij}}. (\#eq:tsne-obj)
\end{equation}

We now circle back to Equation \@ref(eq:pji) and discuss $\sigma_i$. The variance of the Gaussian is allowed to vary depending on where the point is in the space; specifically, it should reflect the density or sparsness of surrounding points. To train this parameter, t-SNE introduces a hyperparameter, *perplexity*, which is a scalar value set by the user, and is defined as

\begin{equation}
Perp(P_i) = 2^{H(P_i)},
\end{equation}
where $H(P_i)$ denotes the entropy of $P_i$,

\begin{equation}
H(P_i) = -\sum_j p_{j|i}\log_2 p_{j|i},
\end{equation}

and $P_i$ is the Gaussian parameterized by $x_i$ and $\sigma_i$. During training, t-SNE looks for $P_i$ that have perpelxity corresponding to the user input. Finally, the objective as stated in Equation \@ref(eq:tsne-obj) is differentiable, and is trained via gradient descent.

While t-SNE is a powerful visualization technique, there are a few points the modeler should be mindful of in practice. We refer the reader to @wattenberg2016how for a more thorough discussion, but stress here that the resulting plot can be very sensitive to the perplexity hyperparameter. In addition, for our use case, which is to explore the learned embeddings, we have fewer data points than most other applications in the literature. Accordingly, our reasonable range of hyperparameter values need to be adjusted down. In Figures \@ref(fig:flood-zone-tsne), we display the t-SNE of the learned embeddings for `flood_zone` in Model 3 with various perplexity settings: 2, 3, 5, and 10; the learning rate and the number of steps are fixed at 100 and 10,000, respectively, for all plots. In the first three plots, we can see the semblance of clusters forming. In particular, flood zone D appears in its own cluster, which implies it is far away from other levels in the embedding space. When the perplexity is set to 10 (and beyond), however, we start to lose the geometric properties.

@@

Compared to the models presented in the previous section, the models presented here are closer to state of the art models that might be used for tabular data modeling problems. To the best of our knowledge, this is the first analysis of these architectures within the actuarial literature.

@@

The final model we consider in this subsection is the TabNet architecture. The structure of the model is quite similar to Model 6, except that a more specialized form of self-attention is used to select the most important covariates for the modeling problem.
